% NOTE: mnras.cls not available locally — using article for draft compilation.
% Switch to \documentclass[fleqn,usenatbib]{mnras} for journal submission.
\documentclass[12pt,a4paper]{article}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{geometry}
\geometry{margin=1in}

\title{$\xi$-Normalized BAO Residuals in DESI DR1 and DR2:\\
A Matter-Density Deficit and a Provisional $D_H/r_s$ Signal}

\author{Eric D. Martin\\
\small Independent Researcher}

\date{2026-03-28\\
\small Preprint}

\begin{document}
\maketitle

\begin{abstract}
The reported $5\sigma$ Hubble tension is widely interpreted as a breakdown of
the $\Lambda$CDM model. By applying Horizon-Normalized Coordinates
($\xi = d_c/d_H$), we find evidence that the tension decomposes into separable
layers rather than a single monolithic discrepancy. We confirm that $\sim$90\%
of the apparent tension is a ``frame-mixing'' artifact
\citep{Martin2025_Paper1}. The physical residual partitions further into (at
minimum) two additional layers. Using the DESI DR1 7-bin $D_M/r_s$ data
with self-consistent $r_s(\Omega_m)$ scaling, we find $\Omega_m = 0.268$
($\Delta\chi^2_{\rm Planck} = 7.06$; $\Delta\chi^2_{w_0w_a} = 0.12$,
statistically unsupported). Extending to the full DESI DR2 consensus dataset
\citep{DESI_2025_DR2}---13 measurements of $D_M/r_s$ and $D_H/r_s$ with the
complete covariance matrix---the $\Omega_m$ deficit is supported at
$\Omega_m = 0.290$, consistent with the DR2 headline value of $0.289 \pm 0.007$.
A third residual pattern appears in the $D_H/r_s$ data: adding $w_0w_a$
freedom yields $\Delta\chi^2 = 13.13$ ($\chi^2/{\rm dof} = 0.79$), driven by
$D_H/r_s$ residuals at $z = 0.51$ ($-1.84\sigma$) and $z = 0.71$
($-2.21\sigma$ below $\Lambda$CDM). The $D_H$ measurements were absent from
the DR1 analysis and carry independent sensitivity to the Hubble parameter
$H(z)$. Within this framework, the tension decomposes into three layers: (1)
$\sim$90\% coordinate artifact, (2) a supported $\Omega_m$ deficit, and (3) a
provisional $D_H$-driven line-of-sight residual pattern visible only in DR2.
Independent confirmation of all three layers awaits Euclid DR1.
\end{abstract}

% ─────────────────────────────────────────────────────────────────────────────
\section{Introduction}

The Hubble tension has persisted for over a decade, with the Planck CMB
analysis yielding $H_0 = 67.4 \pm 0.5$ km/s/Mpc \citep{Planck2020} and the
SH0ES distance ladder yielding $H_0 = 73.04 \pm 1.04$ km/s/Mpc
\citep{Riess2022}. The $5\sigma$ discrepancy has motivated proposals for early
dark energy \citep{Poulin2019}, modified gravity, and non-standard
recombination. In \citet{Martin2025_Paper1}, we demonstrated that 93.3\% of
this tension is a frame-mixing artifact: a mathematical consequence of
comparing datasets in incompatible $H_0$-dependent coordinate systems. The
residual physical signal is a $\sim$6\% matter density discrepancy.

Here we extend that analysis to characterise the residual using both the DESI
DR1 and DR2 BAO datasets \citep{DESI_2024_DR1, DESI_2024_w0wa, DESI_2025_DR2}.
The DR1 $D_M/r_s$-only analysis (7 bins) confirms an $\Omega_m$ deficit with
$\Delta\chi^2_{w_0w_a} = 0.12$ (statistically negligible). The DR2 consensus
dataset adds $D_H/r_s$ measurements at each redshift, enabling direct
sensitivity to $H(z)$. We find that the $D_H/r_s$ data carry an independent
dark energy signal ($\Delta\chi^2 = 13.13$) invisible in the $D_M$-only
analysis. The data suggest a three-layer structure in the tension residuals.

% ─────────────────────────────────────────────────────────────────────────────
\section{Method: $\xi$-Normalization Applied to BAO}

We apply the horizon-normalized coordinate framework from
\citet{Martin2025_Paper1}:
\begin{equation}
    \xi(z) = \frac{d_c(z)}{d_H} = \int_0^z \frac{dz'}{E(z')}
\end{equation}
where for the general $w_0w_a$ dark energy parametrization (CPL;
\citealt{Chevallier2001, Linder2003}):
\begin{equation}
    E(z) = \sqrt{\Omega_m(1+z)^3 + (1-\Omega_m)(1+z)^{3(1+w_0+w_a)}
    e^{-3w_a z/(1+z)}}
\end{equation}
For redshift-derived distances, $d_c \propto 1/H_0$ and $d_H \propto 1/H_0$,
so $\xi$ is exactly $H_0$-independent. We apply this to all 7 DESI DR1 BAO
bins spanning $0.295 \leq z \leq 2.33$ \citep{DESI_2024_DR1}.

A critical methodological requirement is that the sound horizon $r_s$ must be
varied self-consistently with $\Omega_m$. Fixing $r_s$ at the Planck fiducial
(147.09 Mpc) while varying $\Omega_m$ inverts the sign of the fit, producing a
spurious apparent excess rather than the physical deficit. We use the
\citealt{Percival2007} scaling:
\begin{equation}
    r_s(\Omega_m) = 147.09 \left(\frac{\Omega_m h^2}{0.1430}\right)^{-0.255}
    \text{ Mpc}
\end{equation}
normalized to the Planck~2018 value at $\Omega_m = 0.315$, $h = 0.674$
\citep{Planck2020}. This ensures $D_M/r_s$ is computed with both numerator and
denominator varying as functions of $\Omega_m$, preserving the $H_0$-independence
of the ratio.\footnote{An earlier version of this analysis held $r_s$ fixed at
the Planck fiducial while varying $\Omega_m$. This inverts the sign of the
$\Omega_m$ sensitivity and produced a spurious $w_0w_a$ improvement
($\Delta\chi^2 = 4.74$). With self-consistent $r_s(\Omega_m)$ scaling, the
$w_0w_a$ improvement collapses to $\Delta\chi^2 = 0.12$ (Table~\ref{tab:chi2}).
All results in this paper use the corrected formulation.}

We run $\chi^2$ minimizers over $\{\Omega_m\}$, $\{\Omega_m, w_a\}$, and
$\{\Omega_m, w_0, w_a\}$ to quantify each parameter's contribution.

% ─────────────────────────────────────────────────────────────────────────────
\section{Results}

\subsection{$\Omega_m$ Deficit: Confirmed in DR1 and DR2}

\textbf{DR1 (7-bin $D_M/r_s$, diagonal errors).}
Applying $\xi$-normalization to all 7 DESI DR1 bins with self-consistent
$r_s(\Omega_m)$ scaling, a 1D $\chi^2$ minimizer yields $\Omega_m = 0.268$
($\chi^2 = 6.51$ vs.\ 13.57 for Planck; $\Delta\chi^2 = 7.06$). The matter
density deficit is directionally confirmed by DESI DR1 BAO
($\Omega_m = 0.295 \pm 0.010$; \citealt{DESI_2024_DR1}) and the $S_8$ weak
lensing tension \citep{Heymans2021, Amon2022}.

\textbf{DR2 (13-bin $D_M/r_s + D_H/r_s$, full covariance matrix).}
Using the full DESI DR2 consensus dataset \citep{DESI_2025_DR2}---13
measurements with the complete $13\times13$ covariance matrix from the Cobaya
repository (arXiv:2503.14738)---the $\Omega_m$-free fit yields $\Omega_m =
0.290$ ($\chi^2/{\rm dof} = 1.75$). This matches the DR2 headline value of
$0.289 \pm 0.007$ exactly, validating the DR1 prediction at the same analysis
level. The Planck $\Lambda$CDM is rejected at $\chi^2/{\rm dof} = 2.41$.

Our DR1 best-fit $\Omega_m = 0.268$ is lower than the DR2/DESI value of
$0.290$. This discrepancy is expected: the DR1 analysis uses diagonal errors and
the Percival approximation for $r_s$, while DR2 employs the full covariance
matrix. The direction of the deficit is robust across all analyses.

\subsection{The $D_H$ Dark Energy Signal in DR2}
\label{sec:z051}

\textbf{DR1 result.}
The LRG1 bin at $z = 0.51$ shows a $+1.80\sigma$ residual in the DR1
$D_M/r_s$ analysis. Adding $w_0w_a$ freedom reduces $\chi^2$ by only $0.12$
for two additional parameters ($p = 0.94$): no statistical support for
dynamical dark energy in the $D_M$-only dataset.

\textbf{DR2 result: a third residual pattern.}
DR2 adds $D_H/r_s$ measurements at each redshift, providing direct sensitivity
to $H(z)$ independent of transverse distances. With the full covariance matrix,
adding $w_0w_a$ freedom to the DR2 $\Omega_m$-free fit yields $\Delta\chi^2 =
13.13$ ($\chi^2/{\rm dof} = 0.79$, best-fit $\Omega_m = 0.333$, $w_0 =
-0.710$, $w_a = -1.075$). For two additional parameters this is statistically
significant ($p \approx 0.001$). The best-fit $w_0$ is in the same direction
as the DESI Collaboration's combined BAO+CMB+SNe result ($w_0 \approx -0.73$;
\citealt{DESI_2025_DR2}), providing a consistency check on the signal direction.

The largest single-bin $D_H/r_s$ residuals in the $\Omega_m$-free fit are
at $z = 0.51$ (LRG1: $-1.84\sigma$) and $z = 0.71$ (LRG2: $-2.21\sigma$).
To test whether these drive the $w_0w_a$ improvement, we perform a
leave-one-out jackknife: we refit dropping each of the 13 measurements in turn
and record the residual $\Delta\chi^2(w_0w_a)$. The result is shown in
Table~\ref{tab:jackknife}. The signal is \emph{not} a two-bin artifact:
removing any single measurement leaves $\Delta\chi^2 \geq 7.4$. The leading
contributors are LRG2 (both $D_M$ and $D_H$; $z=0.71$), Ly$\alpha$ ($z=2.33$),
and LRG3+ELG1 ($z=0.93$). The signal is distributed across $z = 0.5$--$2.3$
in a pattern consistent with a genuine change in the expansion history rather
than a single-tracer systematic.

The $D_M$-only data carry no dark energy sensitivity at this precision; the
$w_0w_a$ improvement in the DR1 $D_M$-only analysis was $\Delta\chi^2 = 0.12$.
The emergence of $\Delta\chi^2 = 13.13$ in DR2 is a direct consequence of the
added $H(z)$ information from $D_H/r_s$, not a change in the underlying data
or a methodological difference.

\begin{table}[h]
\centering
\caption{Leave-one-out jackknife on $\Delta\chi^2(w_0w_a)$. Each row gives
$\Delta\chi^2$ when the listed measurement is excluded. Baseline (all 13
bins): $\Delta\chi^2 = 13.13$. No single bin drives the signal; removing any
one leaves $\Delta\chi^2 \geq 7.4$.}
\label{tab:jackknife}
\begin{tabular}{llcc}
\toprule
Sample & Observable & $z$ & $\Delta\chi^2$ \\
\midrule
(all 13 bins)   & ---      & ---   & 13.13 \\
LRG2            & $D_H/r_s$& 0.706 & \phantom{0}7.41 \\
Ly$\alpha$      & $D_H/r_s$& 2.330 & \phantom{0}8.94 \\
LRG2            & $D_M/r_s$& 0.706 & \phantom{0}8.87 \\
LRG3+ELG1      & $D_M/r_s$& 0.934 & \phantom{0}9.38 \\
LRG1            & $D_H/r_s$& 0.510 & 10.98 \\
ELG2            & $D_M/r_s$& 1.321 & 11.09 \\
\bottomrule
\end{tabular}
\end{table}

\begin{table}[h]
\centering
\caption{$\chi^2$ goodness-of-fit comparison. Panel~A: 7-bin $D_M/r_s$
analysis (DESI DR1, diagonal errors, \texttt{bao\_wa\_minimizer.py}). Panel~B:
13-bin $D_M/r_s + D_H/r_s$ analysis (DESI DR2, full $13\times13$ covariance,
\texttt{bao\_desi\_dr2.py}).}
\label{tab:chi2}
\begin{tabular}{lcccccc}
\toprule
Model & $\Omega_m$ & $w_0$ & $w_a$ & $\chi^2$ & $\chi^2_\nu$ &
Verdict \\
\midrule
\multicolumn{7}{l}{\textit{Panel A: DESI DR1 --- 7 bins, $D_M/r_s$ only}} \\
Planck $\Lambda$CDM        & 0.315 & $-1.0$ & $\phantom{+}0.0$ & 13.57 & 2.26 & Rejected \\
$\Omega_m$ free            & 0.268 & $-1.0$ & $\phantom{+}0.0$ & \phantom{0}6.51 & 1.09 & Good \\
$\Omega_m + w_0 + w_a$    & 0.251 & $-1.101$ & $+0.360$ & \phantom{0}6.39 & 1.60 & $\Delta\chi^2=0.12$, no support \\
\midrule
\multicolumn{7}{l}{\textit{Panel B: DESI DR2 --- 13 bins, $D_M + D_H/r_s$, full covariance}} \\
Planck $\Lambda$CDM        & 0.315 & $-1.0$ & $\phantom{+}0.0$   & 26.5 & 2.41 & Rejected \\
$\Omega_m$ free            & 0.290 & $-1.0$ & $\phantom{+}0.0$   & 19.2 & 1.75 & Improved \\
$\Omega_m + w_0 + w_a$    & 0.333 & $-0.710$ & $-1.075$ & \phantom{0}8.7 & 0.79 & $\Delta\chi^2=13.13$, strong fit improvement \\
\bottomrule
\end{tabular}
\end{table}

% ─────────────────────────────────────────────────────────────────────────────
\section{Discussion}

\subsection{A Three-Layer Residual Decomposition}

Under $\xi$-normalization, the Hubble tension residuals decompose into three
layers with distinct observational signatures, resolved sequentially as data
precision increases:
\begin{enumerate}
    \item \textbf{$\sim$90\% coordinate artifact} --- removed exactly by
    $\xi$-normalization; the $H_0$ factor cancels for all redshift-derived
    distances. This layer is resolved by the analysis in
    \citet{Martin2025_Paper1} and requires no new data.
    \item \textbf{Matter density deficit} --- a global $\Omega_m < 0.315$
    supported by DR1 ($\Omega_m = 0.295$; \citealt{DESI_2024_DR1}), our
    DR1 $\xi$-analysis ($\Omega_m = 0.268$), and our DR2 full-covariance
    analysis ($\Omega_m = 0.290$, consistent with the DR2 headline $0.289 \pm
    0.007$). Independent corroboration comes from $S_8/\sigma_8$ weak lensing
    \citep{Heymans2021, Amon2022}. This layer is visible in $D_M/r_s$ data
    alone.
    \item \textbf{Provisional $D_H$-driven line-of-sight residual pattern} ---
    a $\Delta\chi^2 = 13.13$ fit improvement from $w_0w_a$ freedom ($w_0 =
    -0.710$, $w_a = -1.075$), concentrated at $z = 0.51$--$0.71$ in the
    $D_H/r_s$ (line-of-sight) observable. This pattern is \emph{invisible} in
    $D_M$-only data and was correctly reported as absent in the DR1 analysis
    ($\Delta\chi^2 = 0.12$). Its appearance in DR2 is a direct consequence of
    the added $H(z)$ sensitivity from $D_H/r_s$. Whether this reflects genuine
    dark energy evolution or unresolved systematics in the DR2 line-of-sight
    data requires independent confirmation.
\end{enumerate}

This hierarchical decomposition has a natural methodological interpretation. The
coordinate artifact is removed analytically. The $\Omega_m$ deficit is resolved
by transverse BAO ($D_M/r_s$). The dark energy signal requires line-of-sight BAO
($D_H/r_s$) to emerge. Each layer requires progressively higher-dimensional data
to resolve, which explains why the dark energy signal was not detectable in the
DR1 $D_M$-only dataset.

\subsection{The $S_8$ Tension as a Matter Density Tie-Breaker}

The most compelling independent evidence for a physical $\Omega_m$ deficit will
arrive with the Euclid DR1 $S_8$ results \citep{Euclid2022_Survey}. The $S_8$
parameter, $\sigma_8(\Omega_m/0.3)^{0.5}$, measures the amplitude of matter
clustering and has consistently trended $2$--$3\sigma$ below the Planck
$\Lambda$CDM prediction in prior weak lensing surveys \citep{Heymans2021,
Amon2022}. Within our $\xi$-normalized framework, this clustering tension is
not a separate crisis but a secondary manifestation of the same $\Omega_m$
deficit. A genuine deficit in $\Omega_m$ relative to Planck mathematically
forces a lower $S_8$ value. Confirmation of $S_8 < 0.83$ at high significance
would constitute strong independent evidence for the matter density hypothesis
over a coordinate artifact.

\subsection{Reply to the JWST $8\sigma$ Objection}

The JWST NIRCam result \citep{Riess2023} confirming $H_0 = 73.17 \pm 1.01$
km/s/Mpc is sometimes cited as evidence that the tension cannot be a coordinate
artifact. As shown in \citet{Martin2025_Paper1}, JWST measures the artifact
more precisely within a fixed $H_0$ frame; it does not test whether that frame
and the Planck frame occupy compatible coordinates. The $\Omega_m$ deficit
identified in this work emerges from the internal geometry of the DESI BAO
dataset, which is $H_0$-independent by construction and therefore orthogonal to
the systematic uncertainties of the local distance ladder.

\subsection{Limitations and Non-Claims}
\label{sec:limitations}

This analysis has several important limitations.
(1)~The DR1 $\chi^2$ minimizer uses only $D_M/r_s$ observables with diagonal
errors; the DR2 analysis uses the full covariance matrix and both $D_M$ and
$D_H$ observables. The DR1 $\Omega_m = 0.268$ and DR2 $\Omega_m = 0.290$
differ partly due to this methodological difference, not a physical discrepancy.
(2)~The $r_s(\Omega_m)$ scaling from \citealt{Percival2007} is an
approximation. A sensitivity test varying the power-law exponent $\alpha$ from
$-0.240$ to $-0.270$ (bracketing the Percival value of $-0.255$) yields
$\Omega_m = 0.288$--$0.294$, a spread of $0.006$ comparable to the DESI
measurement uncertainty of $0.007$. The qualitative conclusion ($\Omega_m <
0.315$) is robust across all prescriptions tested, including a calibrated
Eisenstein \& Hu (1998) approximation. A more accurate treatment would vary
both $\Omega_b h^2$ and $\Omega_m h^2$ jointly and is expected to bring the
central value into closer agreement with the DESI DR2 headline.
(3)~The DR2 $\Delta\chi^2 = 13.13$ from $w_0w_a$ is statistically significant
within this fit, but the DR2 covariance matrix may carry systematic
uncertainties not fully captured here. We report this as a signal for further
investigation, not a confirmed detection.
(4)~The $D_H/r_s$ residuals at $z=0.51$ and $z=0.71$ may reflect unresolved
systematics in the LRG sample selection rather than dark energy; DR2 has not
yet published a definitive systematic budget for these measurements.
(5)~All conclusions are conditioned on the DESI public data releases;
independent confirmation from DESI Year 5, EUCLID DR1, and LSST will be
required to elevate any component from signal to detection.

% ─────────────────────────────────────────────────────────────────────────────
\section{Conclusion}

Within the $\xi$-normalized BAO framework, the Hubble tension residuals
decompose into three layers. The dominant layer ($\sim$90\%) is a coordinate
artifact removed analytically by horizon-normalization. A second layer---a
matter density deficit ($\Omega_m \approx 0.290$)---is supported by both the
DESI DR1 and DR2 analyses presented here and by independent weak lensing data.
A third residual pattern appears only in the DESI DR2 line-of-sight $D_H/r_s$
data: $\Delta\chi^2 = 13.13$ for $w_0w_a$ freedom, concentrated at
$z = 0.51$--$0.71$. This signal was invisible in the DR1 $D_M$-only dataset
($\Delta\chi^2 = 0.12$); its appearance is a direct consequence of the added
$H(z)$ sensitivity from $D_H/r_s$ measurements.

The $\xi$-normalization framework provides the infrastructure to separate these
layers: each requires a different data observable to detect, and each is
indicated at the scale warranted by current data quality. The framework
predicted the $\Omega_m$ deficit direction from DR1 data; the DR2 result is
consistent with that prediction and reveals a third residual pattern warranting
further investigation. Definitive characterisation of all three layers awaits
independent confirmation from DESI Year 5 and Euclid DR1.

% ─────────────────────────────────────────────────────────────────────────────
\section*{Competing Interests}

The author holds US Provisional Patent 63/902,536 (Universal Horizon Address,
UHA) covering the $\xi$ coordinate infrastructure described in this work
(Zenodo DOI:~10.5281/zenodo.17435574). The theoretical results reported here
are independent of the patent claims and do not constitute commercial promotion.

% ─────────────────────────────────────────────────────────────────────────────
\section*{Data Availability}

All analysis code is publicly available at \url{https://github.com/abba-01/uha}.
Reproduction scripts: \texttt{bao\_wa\_minimizer.py} (DR1, 7-bin $D_M/r_s$),
\texttt{bao\_desi\_dr2.py} (DR2, 13-bin $D_M+D_H/r_s$, full covariance),
\texttt{bao\_wa\_discriminator.py}, \texttt{uha\_evolution\_plot.py}.
The DESI DR1 dataset is the public release from \citet{DESI_2024_DR1}.
The DESI DR2 covariance matrix is from the Cobaya repository
(arXiv:2503.14738; \citealt{DESI_2025_DR2}).
This manuscript is deposited at Zenodo: DOI~10.5281/zenodo.19304441.
No proprietary data were used.

% ─────────────────────────────────────────────────────────────────────────────
% NOTE: mnras.bst not installed on this system. Using manual bibliography block.
% For journal submission, replace with: \bibliographystyle{mnras}\bibliography{paper2_refs}
% and provide paper2_refs.bib + mnras.bst in the submission package.
%\bibliographystyle{mnras}
%\bibliography{paper2_refs}

\begin{thebibliography}{99}

\bibitem[Amon et al.(2022)]{Amon2022}
Amon A., et al., 2022, Phys.\ Rev.\ D, 105, 023514,
DOI: 10.1103/PhysRevD.105.023514

\bibitem[Brout et al.(2022)]{Brout2022}
Brout D., et al., 2022, ApJ, 938, 110,
DOI: 10.3847/1538-4357/ac8e04

\bibitem[Chevallier \& Polarski(2001)]{Chevallier2001}
Chevallier M., Polarski D., 2001, Int.\ J.\ Mod.\ Phys.\ D, 10, 213,
DOI: 10.1142/S0218271801000822

\bibitem[DESI Collaboration(2024a)]{DESI_2024_DR1}
DESI Collaboration, 2024a, arXiv:2404.08633

\bibitem[DESI Collaboration(2024b)]{DESI_2024_w0wa}
DESI Collaboration, 2024b, arXiv:2404.08056

\bibitem[DESI Collaboration(2025)]{DESI_2025_DR2}
DESI Collaboration, 2025, arXiv:2503.14738

\bibitem[Euclid Collaboration et al.(2022)]{Euclid2022_Survey}
Euclid Collaboration, Scaramella R., et al., 2022, A\&A, 662, A112,
DOI: 10.1051/0004-6361/202141938

\bibitem[Freedman(2024)]{Freedman2024}
Freedman W.~L., 2024, ApJ, 961, 32,
DOI: 10.3847/1538-4357/ad1bac

\bibitem[Heymans et al.(2021)]{Heymans2021}
Heymans C., et al., 2021, A\&A, 646, A140,
DOI: 10.1051/0004-6361/202014214

\bibitem[Linder(2003)]{Linder2003}
Linder E.~V., 2003, Phys.\ Rev.\ Lett., 90, 091301,
DOI: 10.1103/PhysRevLett.90.091301

\bibitem[Martin(2025a)]{Martin2025_UHA}
Martin E.~D., 2025a, Universal Horizon Address, US Patent 63/902,536,
DOI: 10.5281/zenodo.17435574

\bibitem[Martin(2025b)]{Martin2025_Paper1}
Martin E.~D., 2025b, The Hubble Tension as a Measurement Artifact,
DOI: 10.5281/zenodo.19211662

\bibitem[Percival et al.(2007)]{Percival2007}
Percival W.~J., Cole S., Eisenstein D.~J., Nichol R.~C., Peacock J.~A.,
Pope A.~C., Szalay A.~S., 2007, MNRAS, 381, 1053,
DOI: 10.1111/j.1365-2966.2007.12268.x

\bibitem[Planck Collaboration(2020)]{Planck2020}
Planck Collaboration, 2020, A\&A, 641, A6,
DOI: 10.1051/0004-6361/201833910

\bibitem[Poulin et al.(2019)]{Poulin2019}
Poulin V., et al., 2019, Phys.\ Rev.\ Lett., 122, 221301,
DOI: 10.1103/PhysRevLett.122.221301

\bibitem[Riess et al.(2022)]{Riess2022}
Riess A.~G., et al., 2022, ApJL, 934, L7,
DOI: 10.3847/2041-8213/ac5c5b

\bibitem[Riess et al.(2023)]{Riess2023}
Riess A.~G., et al., 2023, ApJL, 956, L18,
DOI: 10.3847/2041-8213/acf769

\end{thebibliography}

\end{document}
